### 2019-12-27 Statistical Methods/08_Regression.rmd
statistical methods
statistical methods

>>>>

Maximal oxygen uptake (MOU) is a measure of an individual's physical fitness. You want to know how MOU is related to how fast someone can run a mile. Suppose you randomly sample six athletes and get:

MOU (kl/kg): 63.3, 60.1, 53.6, 58.8, 67.5, 62.5
Time (seconds): 241.5, 249.8, 246.1, 232.4, 237.2, 238.4

Compute the least squres regression line and comment on what the results suggest.

```{r, echo = T, fig.width=8, fig.height=3.5}
dat <- data.table(
   MOU = c(63.3, 60.1, 53.6, 58.8, 67.5, 62.5),
   Time = c(241.5, 249.8, 246.1, 232.4, 237.2, 238.4)
)

fit <- lm(Time ~ MOU, data = dat)

ggplot(dat, aes(MOU, Time)) +
   geom_point() +
   geom_smooth(method = "lm") +
   labs(title = "Time vs. MOU")

Generally, time decreases as MOU increases.

### 8.11

Verify that for the following pairs of points, the least squares regression line has a slope of zero. Plot the points and comment on the assumption that the regression line is straight.

X: 1, 2, 3, 4, 5, 6
Y: 1, 4, 7, 7, 4, 1

```{r, echo = T, fig.width=8, fig.height=3.5}
dat <- data.table(
   X = c(1, 2, 3, 4, 5, 6),
   Y = c(1, 4, 7, 7, 4, 1)
)

fit <- lm(Y ~ X, data = dat)

coef(fit)
```

### 8.12

Repeat the last exercise, only for points:

X: 1, 2, 3, 4, 5, 6
Y: 4, 5, 6, 7, 8, 2

```{r, echo = T, fig.width=8, fig.height=3.5}
dat <- data.table(
   X = c(1, 2, 3, 4, 5, 6),
   Y = c(4, 5, 6, 7, 8, 2)
)

fit <- lm(Y ~ X, data = dat)

coef(fit)
```

### 8.13

Vitamin A is required for good health. However, one bite of polar bear liver results in death becasue it contains a high concentration of vitamin A.

Comment on this fact in terms of extrapolation.

_Vitaim A as a function of health is a bounded variable. Predicting a response when the independent variables are far from what has been observed can lead to undefined results (death)._

### 8.14

Socket et al. (1987) report data related to patterns of residual insulin secretion in children. A portion of the study was concerned with whether age can be used to predict the logarithm of C-peptide concentrations at diagnosis. The observed values are (data file):

Replicate the _LOESS_ smoothed curve in fig 8.4.1.

```{r, echo = T, fig.width=8, fig.height=3.5}
dat <- data.table::fread(paste0(data.dir, "diabetes_sockett_dat.txt"), fill = T)

ggplot(dat, aes(age, pep)) +
   geom_point() +
   geom_smooth(method = "loess")

### 8.15

### 8.16
### 8.17
### 8.18
### 8.19
### 8.20
